Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-11-07 paper/paper-MinD.tex
changes to paper for page limits Tested everything on crystal.

\keywords{\fixme{E. Coli | MinD | oscillation | stochastic | deterministic | simulation}}

\abbreviations{\fixme{SAM, self-assembled monolayer; OTS,
octadecyltrichlorosilane}}

\dropcap{I}t is vital that during the process of bacterial cell division a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  Instrumental to this process in \emph{Escherichia coli} is a
long FtsZ polymer chain that develops on the cell wall in the center
region of the cell, helping dictate the plane of
division~\cite{adams2009bacterial, lutkenhaus2007assembly}. Previous
experimental studies have shown that the MinC protein, known to
inhibit the FtZ polymer~\cite{shen2010examination}, exhibits regular
pole to pole oscillatory behavior between both ends of the wild-type
pill-shaped cell.  It thus has a higher time averaged concentration in
the cell poles than in the center region, which aides in preventing
the FtZ from developing in the wrong region.  The MinC is recruited to
these poles by MinD, which itself interacts with another protein,
MinE, in a system which exhibits pole-to-pole oscillatory
behavoir~\cite{hu1999topological, fu2001mine, shapiro2009and,
  yu1999ftsz, raskin1999rapid, meacci2005min, raskin1999minde}.

Previous experimental studies have shown that the MinD protein system
is capable of exhibiting oscillations in round
shapes~\cite{fange2006noise} as well as in connected three pronged
tube shapes~\cite{varma2008min}, in which the oscillations seem to
seek out the extreme poles in the
cell~\cite{juarez2010changes,corbin2002exploring}.  One conclusion of
these studies has been that the MinD system robustly forms bipolar
oscillations regardless of variations in cellular shape.

@@

\fixme{look for easily cut citations.}

@@

In this paper we use Huang's 2003 model~\cite{huang2003dynamic} in
both deterministic and stochastic variants to study flattened cells
similar to those observed by Mannik~\cite{mannik2009bacterial}.

>>>>

\keywords{MinD | \emph{Escherichia coli} | stochastic | deterministic }

%% \abbreviations{SAM, self-assembled monolayer; OTS,
%% octadecyltrichlorosilane}

\dropcap{I}t is vital that during the process of bacterial cell
division a cell avoid minicelling, or splitting into daughter cells
with lopsided volumes.  Instrumental to this process in
\emph{Escherichia coli} is a long FtsZ polymer chain that develops on
the cell wall in the center region of the cell, helping dictate the
plane of division~\cite{adams2009bacterial,
  lutkenhaus2007assembly}. Previous experimental studies have shown
that the MinC protein, known to inhibit the FtZ
polymer~\cite{shen2010examination}, exhibits regular pole to pole
oscillatory behavior between both ends of the wild-type pill-shaped
cell.  It thus has a higher time averaged concentration in the cell
poles than in the center region, which aides in preventing the FtZ
from developing in the wrong region.  The MinC is recruited to these
poles by MinD, which itself interacts with another protein, MinE, in a
system which exhibits pole-to-pole oscillatory
behavoir~\cite{shapiro2009and, yu1999ftsz,
  meacci2005min,huang2003dynamic,kerr2006division,mannik2009bacterial}.
%% \cite{hu1999topological, fu2001mine, shapiro2009and,
%%   yu1999ftsz, raskin1999rapid, meacci2005min, raskin1999minde}.

Previous experimental studies have shown that the MinD protein system
is capable of exhibiting oscillations in round
shapes~\cite{fange2006noise} as well as in connected three pronged
tube shapes~\cite{varma2008min}, in which the oscillations seem to
seek out the extreme poles in the cell.  One conclusion of these
studies has been that the MinD system robustly forms bipolar
oscillations regardless of variations in cellular shape.

@@

@@

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's 2003 model~\cite{huang2003dynamic} in
both deterministic and stochastic variants to study flattened cells
similar to those observed by Mannik~\cite{mannik2009bacterial}.
